---
title: "Hierarchical IGT models"
author: "Line Kruse"
date: "3/26/2020"
output: html_document
---

```{r setup, include=FALSE}
set.seed(1982)
library(pacman)
p_load(R2jags)

setwd("~/Documents/Advanced Cognitive Modeling /Exercise 4 - Fitting data and hierarchical models")

```

#Load and clean data for PVL Delta and ORL model 

```{r}
HC_data <-read.delim("~/Documents/Advanced Cognitive Modeling /Exercise 4 - Fitting data and hierarchical models/rawData/IGTdata_healthy_control.txt")

#Long to wide format 
subIDs <- unique(HC_data$subjID)
nsubs <- length(subIDs)
ntrials_max <- 100

x_raw <- HC_data$deck #choices raw
X_raw <- HC_data$gain + HC_data$loss #rewards (loss and rewards are tracked together in the PVL delta model)

###Assign choices and rewards in a trial*subject matrix 
#We have different numbers of trials for each subject - we fix this by padding arrays < 100 
#We also need to record number of valid trials for each subject (no NAs)
#Then we run the JAGS model on the valid trials only 

ntrials_all <- array(0, c(nsubs))
x_all <- array(0, c(nsubs, ntrials_max))
X_all <- array(0, c(nsubs, ntrials_max))

for (s in 1:nsubs){
  
  #Record n trials for subject s
  ntrials_all[s] <- length(x_raw[HC_data$subjID==subIDs[s]])
  
  #Pad trials with NA if n trials < maximum (100)
  x_sub <- x_raw[HC_data$subjID==subIDs[s]]
  length(x_sub) <- ntrials_max
  
  X_sub <- X_raw[HC_data$subjID==subIDs[s]]
  length(X_sub) <- ntrials_max
  
  #Assign arrays 
  x_all[s,] <- x_sub 
  X_all[s,] <- X_sub
  
}

X_all <- X_all/100


```

#Clean data for VSE model 

```{r}
#Long to wide format 
subIDs <- unique(HC_data$subjID)
nsubs <- length(subIDs)
ntrials_max <- 100

x_raw <- HC_data$deck #choices raw
r_raw <- HC_data$gain #rewards 
l_raw <- HC_data$loss #loss 
#In VSE model, rewards and losses are calculated separately 

###Assign choices and rewards in a trial*subject matrix 
#We have different numbers of trials for each subject - we fix this by padding arrays < 100 
#We also need to record number of valid trials for each subject (no NAs)
#Then we run the JAGS model on the valid trials only 

ntrials_all <- array(0, c(nsubs))
x_all <- array(0, c(nsubs, ntrials_max))
r_all <- array(0, c(nsubs, ntrials_max))
l_all <- array(0, c(nsubs, ntrials_max))

for (s in 1:nsubs){
  
  #Record n trials for subject s
  ntrials_all[s] <- length(x_raw[HC_data$subjID==subIDs[s]])
  
  #Pad trials with NA if n trials < maximum (100)
  x_sub <- x_raw[HC_data$subjID==subIDs[s]]
  length(x_sub) <- ntrials_max
  
  r_sub <- r_raw[HC_data$subjID==subIDs[s]]
  length(r_sub) <- ntrials_max
  
  l_sub <- l_raw[HC_data$subjID==subIDs[s]]
  length(l_sub) <- ntrials_max
  
  #Assign arrays 
  x_all[s,] <- x_sub 
  r_all[s,] <- r_sub
  l_all[s,] <- l_sub
  
}

r_all <- r_all/100
l_all <- l_all/100
l_all <- abs(l_all)

```



#Fit hierarchical jags model to the healthy control data - PVL Delta 

```{r}
x <- x_all #choices
X <- X_all #rewards 
ntrials <- ntrials_all

data <- list("x", "X", "ntrials", "nsubs")
params <- c("mu_w", "mu_A", "mu_theta", "mu_a", "lambda_w", "lambda_A", "lambda_theta", "lambda_a")

samples <- jags.parallel(data, inits = NULL, params, 
                    model.file = "Hierarchical_PVL_DeltaModel_jags.txt", 
                    n.chains =3, n.iter =5000, n.burnin = 1000, n.thin = 1)

```

#Plot posterior distributions from hierarchical model - PVL Delta 

```{r}
#Extract posterior distribution
mu_w_post <- samples$BUGSoutput$sims.list$mu_w
mu_A_post <- samples$BUGSoutput$sims.list$mu_A
mu_theta_post <- samples$BUGSoutput$sims.list$mu_theta
mu_a_post <- samples$BUGSoutput$sims.list$mu_a

#Plot posterior distributions 
png("HiraFitPVL_Delta.png") 
par(mfrow=c(2,2))
plot(density(mu_w_post))
plot(density(mu_A_post))
plot(density(mu_theta_post))
plot(density(mu_a_post))
dev.off()

#Extract MAP 
pMean_w <- density(mu_w_post)$x[which(density(mu_w_post)$y==max(density(mu_w_post)$y))]
pMean_A <- density(mu_A_post)$x[which(density(mu_A_post)$y==max(density(mu_A_post)$y))]
pMean_theta <- density(mu_theta_post)$x[which(density(mu_theta_post)$y==max(density(mu_theta_post)$y))]
pMean_a <- density(mu_a_post)$x[which(density(mu_a_post)$y==max(density(mu_a_post)$y))]

```



#Fit hierarchical jags model to the healthy control data - ORL model 


```{r}
x <- x_all
r <- X_all
ntrials <- ntrials_all
  
data <- list("x", "r", "ntrials", "nsubs")
params <- c("mu_Arew", "mu_Apun", "mu_betaF", "mu_betaP", "mu_K", "lambda_Arew", "lambda_Apun", "lambda_betaF", "lambda_betaP", "lambda_K")

samples <- jags.parallel(data, inits = NULL, params, 
                         model.file = "Hier_ORL.txt", 
                         n.chains =3, n.iter =5000, n.burnin = 1000, n.thin = 1)

samples

```

#Plot posterior distributions from hierarchical model - ORL 

```{r}
#Extract posterior distribution
mu_Arew_post <- samples$BUGSoutput$sims.list$mu_Arew
mu_Apun_post <- samples$BUGSoutput$sims.list$mu_Apun
mu_betaF_post <- samples$BUGSoutput$sims.list$mu_betaF
mu_betaP_post <- samples$BUGSoutput$sims.list$mu_betaP
mu_K_post <- samples$BUGSoutput$sims.list$mu_K

#Plot posterior distributions 
png("HiraFit_ORL.png") 
par(mfrow=c(3,2))
plot(density(mu_Arew_post))
plot(density(mu_Apun_post))
plot(density(mu_betaF_post))
plot(density(mu_betaP_post))
plot(density(mu_K_post))
dev.off()
```



#Fit hierarchical jags model to the healthy control data - VSE model 

```{r}
#Fit jags model 
x <- x_all
r <- r_all
l <- l_all
  
ntrials <- ntrials_all
    
#Run parameter inference 
data <- list("x", "r", "l", "ntrials", "nsubs")
params <- c("mu_theta", "mu_delta", "mu_a", "mu_phi", "mu_C", "lambda_theta","lambda_delta", "lambda_a", "lambda_phi", "lambda_C")

samples <- jags.parallel(data, inits = NULL, params, 
                         model.file = "Hier_VSE.txt", 
                         n.chains =3, n.iter =5000, n.burnin = 1000, n.thin = 1)

samples
```

#Plot posterior distributions from hierarchical model - VSE 

```{r}
#Extract posterior distribution
mu_theta_post <- samples$BUGSoutput$sims.list$mu_theta
mu_delta_post <- samples$BUGSoutput$sims.list$mu_delta
mu_a_post <- samples$BUGSoutput$sims.list$mu_a
mu_phi_post <- samples$BUGSoutput$sims.list$mu_phi
mu_C_post <- samples$BUGSoutput$sims.list$mu_C

#Plot posterior distributions 
png("HiraFit_VSE.png") 
par(mfrow=c(3,2))
plot(density(mu_theta_post))
plot(density(mu_delta_post))
plot(density(mu_a_post))
plot(density(mu_phi_post))
plot(density(mu_C_post))
dev.off()
```



