---
title: "Posterior predictive checks"
author: "Line Kruse"
date: "3/25/2020"
output: html_document
---

#Exercise 1 of module 4
#Posterior predictive checks - used to evaluate the descriptive adequacy 
#For the three models of the IGT (from exercise 3)
#The three models are PVL Delta model, ORL model, and VSE model 



```{r setup, include=FALSE}
set.seed(1982)
library(pacman)
p_load(R2jags)

setwd("~/Documents/Advanced Cognitive Modeling /Exercise 4 - Fitting data and hierarchical models/rawData")

#Load healhty control data
HC_data <-read.delim("~/Documents/Advanced Cognitive Modeling /Exercise 4 - Fitting data and hierarchical models/rawData/IGTdata_healthy_control.txt")
```

#Prepare data for PVL Delta and ORL model - healthy controls 

```{r}
#Long to wide format 
subIDs <- unique(HC_data$subjID)
nsubs <- length(subIDs)
ntrials_max <- 100

x_raw <- HC_data$deck #choices raw
X_raw <- HC_data$gain + HC_data$loss #rewards (loss and rewards are tracked together in the PVL delta model)

###Assign choices and rewards in a trial*subject matrix 
#We have different numbers of trials for each subject - we fix this by padding arrays < 100 
#We also need to record number of valid trials for each subject (no NAs)
#Then we run the JAGS model on the valid trials only 

ntrials_all <- array(0, c(nsubs))
x_all <- array(0, c(nsubs, ntrials_max))
X_all <- array(0, c(nsubs, ntrials_max))

for (s in 1:nsubs){
  
  #Record n trials for subject s
  ntrials_all[s] <- length(x_raw[HC_data$subjID==subIDs[s]])
  
  #Pad trials with NA if n trials < maximum (100)
  x_sub <- x_raw[HC_data$subjID==subIDs[s]]
  length(x_sub) <- ntrials_max
  
  X_sub <- X_raw[HC_data$subjID==subIDs[s]]
  length(X_sub) <- ntrials_max
  
  #Assign arrays 
  x_all[s,] <- x_sub 
  X_all[s,] <- X_sub
  
}

```

#Prepare data for VSE model - healthy controls 

```{r}
#Long to wide format 
subIDs <- unique(HC_data$subjID)
nsubs <- length(subIDs)
ntrials_max <- 100

x_raw <- HC_data$deck #choices raw
r_raw <- HC_data$gain #rewards 
l_raw <- HC_data$loss #loss 
#In VSE model, rewards and losses are calculated separately 

###Assign choices and rewards in a trial*subject matrix 
#We have different numbers of trials for each subject - we fix this by padding arrays < 100 
#We also need to record number of valid trials for each subject (no NAs)
#Then we run the JAGS model on the valid trials only 

ntrials_all <- array(0, c(nsubs))
x_all <- array(0, c(nsubs, ntrials_max))
r_all <- array(0, c(nsubs, ntrials_max))
l_all <- array(0, c(nsubs, ntrials_max))

for (s in 1:nsubs){
  
  #Record n trials for subject s
  ntrials_all[s] <- length(x_raw[HC_data$subjID==subIDs[s]])
  
  #Pad trials with NA if n trials < maximum (100)
  x_sub <- x_raw[HC_data$subjID==subIDs[s]]
  length(x_sub) <- ntrials_max
  
  r_sub <- r_raw[HC_data$subjID==subIDs[s]]
  length(r_sub) <- ntrials_max
  
  l_sub <- l_raw[HC_data$subjID==subIDs[s]]
  length(l_sub) <- ntrials_max
  
  #Assign arrays 
  x_all[s,] <- x_sub 
  r_all[s,] <- r_sub
  l_all[s,] <- l_sub
  
}

```


#Apply the model to the data of one subject - PVL Delta 

```{r}
#First, we fit one subject to make sure everything works 
setwd("~/Documents/Advanced Cognitive Modeling /Exercise3 - Iowa Gambling Task") #where we have the models 

#Assign subject 1 to data inputs 
x <- x_all[1,] #choice 
X <- X_all[1,] #reward 

#Set up and run JAGS model on one subject 
data <- list("x", "X", "ntrials")
params <- c("w", "A", "theta", "a", "p")

samples <- jags.parallel(data, inits = NULL, params, 
                         model.file = "PVL_DeltaModel_jags.txt", 
                         n.chains =3, n.iter =5000, n.burnin = 1000, n.thin = 1)


#Plot posterior for parameters 
par(mfrow=c(2,2))
plot(density(samples$BUGSoutput$sims.list$w))
plot(density(samples$BUGSoutput$sims.list$A))
plot(density(samples$BUGSoutput$sims.list$theta))
plot(density(samples$BUGSoutput$sims.list$a))

```

#Posterior prediction for one subject - PVL Delta 

```{r}
#How would you expect the data to look on the basis of the above posteriors? 

#Posterior prediction for p parameter 
#Extract posterior for probability of choosing a given deck, p 
p_post <- samples$BUGSoutput$sims.list$p

#Plot probability of each deck on trial 32 
par(mfrow=c(2,2))
plot(density(p_post[,32,1]))
plot(density(p_post[,32,2])) #Highest probability of deck 2
plot(density(p_post[,32,3]))
plot(density(p_post[,32,4]))

#Which deck was chosen on trial 32
x[32] #Deck 3 was chosen 

#Model guess = deck 2 
#Actual guess = deck 3

#Posterior prediction for all trials (which deck had highest prob at each trials)
x_predict <- array(0, c(ntrials))
for (t in 1:ntrials){
  
  #Infer maximum a posteriori (MAP)
  p_predict <- c(
    density(p_post[,t,1])$x[which(density(p_post[,t,1])$y==max(density(p_post[,t,1])$y))],
    density(p_post[,t,2])$x[which(density(p_post[,t,2])$y==max(density(p_post[,t,2])$y))],
    density(p_post[,t,3])$x[which(density(p_post[,t,3])$y==max(density(p_post[,t,3])$y))],
    density(p_post[,t,4])$x[which(density(p_post[,t,4])$y==max(density(p_post[,t,4])$y))])
  
  x_predict[t] <- which.max(p_predict)
  
}

#Compare predictions to the data - how many was correct? 
sum(x_predict==x)  #47/100 correct predictions 


```


##Apply JAGS model to all subjects and run posterior prediction - PVL Delta 
#####Posterior prediction = descriptive adequacy 

```{r}
#Array for number of each correct prediction for each subject 
pred_success <- array(0, c(nsubs))

setwd("~/Documents/Advanced Cognitive Modeling /Exercise3 - Iowa Gambling Task") #where we have the models 
for (s in 1:nsubs){
  
  #Fit jags model 
  x <- x_all[s,]
  r <- X_all[s,]
  
  ntrials <- ntrials_all[s]
  
  data <- list("x", "X", "ntrials")
  params <- c("w", "A", "theta", "a", "p")

  samples <- jags.parallel(data, inits = NULL, params, 
                         model.file = "PVL_DeltaModel_jags.txt", 
                         n.chains =3, n.iter =5000, n.burnin = 1000, n.thin = 1)
  
  #Extract posterior for p parameter 
  p_post <- samples$BUGSoutput$sims.list$p
  
  #Match predictions to actual choice 
  x_predict <- array(0, c(ntrials))
  for (t in 1:ntrials){
    
    #Infer maximum a posteriori (MAP)
    p_predict <- c(
      density(p_post[,t,1])$x[which(density(p_post[,t,1])$y==max(density(p_post[,t,1])$y))],
      density(p_post[,t,2])$x[which(density(p_post[,t,2])$y==max(density(p_post[,t,2])$y))],
      density(p_post[,t,3])$x[which(density(p_post[,t,3])$y==max(density(p_post[,t,3])$y))],
      density(p_post[,t,4])$x[which(density(p_post[,t,4])$y==max(density(p_post[,t,4])$y))])
  
   x_predict[t] <- which.max(p_predict)
    
  }
  
  #How many trials did the model infer correctly for each subject? 
  x <- x[!is.na(x)]
  pred_success[s] <- sum(x_predict==x, na.rm = TRUE)
  print(s)
  
}

descAdequacy_PVL <- pred_success

```

##Apply JAGS model to all subjects and run posterior prediction - ORL modle 

```{r}
pred_success <- array(0, c(nsubs))

setwd("~/Documents/Advanced Cognitive Modeling /Exercise3 - Iowa Gambling Task") #where we have the models 
for (s in 1:nsubs){
  
  #Fit jags model 
  x <- x_all[s,]
  r <- X_all[s,]
  
  ntrials <- ntrials_all[s]
  
  data <- list("x", "r", "ntrials")
  params <- c("Arew", "Apun", "betaF", "betaP", "K", "p")

  samples <- jags.parallel(data, inits = NULL, params, 
                         model.file = "ORL_Model_jags.txt", 
                         n.chains =3, n.iter =5000, n.burnin = 1000, n.thin = 1)
  
  #Extract posterior for p parameter 
  p_post <- p_post <- samples$BUGSoutput$sims.list$p
  
  #Match predictions to actual choice 
  x_predict <- array(0, c(ntrials))
  for (t in 1:ntrials){
    
    #Infer maximum a posteriori (MAP)
    p_predict <- c(
      density(p_post[,t,1])$x[which(density(p_post[,t,1])$y==max(density(p_post[,t,1])$y))],
      density(p_post[,t,2])$x[which(density(p_post[,t,2])$y==max(density(p_post[,t,2])$y))],
      density(p_post[,t,3])$x[which(density(p_post[,t,3])$y==max(density(p_post[,t,3])$y))],
      density(p_post[,t,4])$x[which(density(p_post[,t,4])$y==max(density(p_post[,t,4])$y))])
  
   x_predict[t] <- which.max(p_predict)
    
  }
  
  #How many trials did the model infer correctly for each subject? 
  x <- x[!is.na(x)]
  pred_success[s] <- sum(x_predict==x, na.rm = TRUE)
  print(s)
  
}

descAdequacy_ORL <- pred_success

```

##Apply JAGS model to all subjects and run posterior prediction - VSE model 

```{r}
pred_success <- array(0, c(nsubs))

setwd("~/Documents/Advanced Cognitive Modeling /Exercise3 - Iowa Gambling Task") #where we have the models 
for (s in 1:nsubs){
  
  #Fit jags model 
  x <- x_all[s,]
  r <- r_all[s,]
  l <- l_all[s,]
  
  ntrials <- ntrials_all[s]
    
  #Run parameter inference 
  data <- list("x", "r", "l", "ntrials")
  params <- c("theta", "delta", "a", "phi", "C", "p")

  samples <- jags.parallel(data, inits = NULL, params, 
                         model.file = "ORL_Model_jags.txt", 
                         n.chains =3, n.iter =5000, n.burnin = 1000, n.thin = 1)
  
  #Extract posterior for p parameter 
  p_post <- p_post <- samples$BUGSoutput$sims.list$p
  
  #Match predictions to actual choice 
  x_predict <- array(0, c(ntrials))
  for (t in 1:ntrials){
    
    #Infer maximum a posteriori (MAP)
    p_predict <- c(
      density(p_post[,t,1])$x[which(density(p_post[,t,1])$y==max(density(p_post[,t,1])$y))],
      density(p_post[,t,2])$x[which(density(p_post[,t,2])$y==max(density(p_post[,t,2])$y))],
      density(p_post[,t,3])$x[which(density(p_post[,t,3])$y==max(density(p_post[,t,3])$y))],
      density(p_post[,t,4])$x[which(density(p_post[,t,4])$y==max(density(p_post[,t,4])$y))])
  
   x_predict[t] <- which.max(p_predict)
    
  }
  
  #How many trials did the model infer correctly for each subject? 
  x <- x[!is.na(x)]
  pred_success[s] <- sum(x_predict==x, na.rm = TRUE)
  print(s)
  
}

descAdequacy_VSE <- pred_success

```

#Compare the descriptive adequacy of the three models 

```{r}
mean(descAdequacy_PVL) #mean accuracy = 37.15
mean(descAdequacy_ORL) #mean accuracy = 40.19
mean(descAdequacy_VSE) #mean accuracy = 37.48
```

